---
title: "Robust standard errors with parglm and the sandwich package and regression tables with gtsummary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Robust standard errors with parglm and the sandwich package and regression tables with gtsummary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  eval = requireNamespace("sandwich", quietly = TRUE) &&
         requireNamespace("lmtest",   quietly = TRUE)
)
```

Since `parglm()` returns a standard `glm` object, it works directly with the
[`sandwich`](https://cran.r-project.org/package=sandwich) package for
heteroskedasticity-consistent (HC) and cluster-robust standard errors.
The [`lmtest`](https://cran.r-project.org/package=lmtest) package provides
`coeftest()` for displaying results with alternative covariance matrices.

## Setup

```{r setup, message = FALSE}
library(parglm)
library(sandwich)
library(lmtest)
```

We simulate a Poisson dataset with 20 clusters of 10 observations each,
where a cluster-level random effect induces within-cluster correlation.

```{r simulate}
set.seed(1)
n          <- 200
cluster_id <- rep(1:20, each = 10)
x1         <- rnorm(n)
x2         <- rnorm(n)
u          <- rep(rnorm(20, sd = 0.5), each = 10)  # cluster random effect
y          <- rpois(n, exp(0.5 + 0.3 * x1 - 0.2 * x2 + u))
dat        <- data.frame(y = y, x1 = x1, x2 = x2, cluster = cluster_id)
```

## Fitting the model

```{r fit}
fit <- parglm(y ~ x1 + x2, data = dat, family = poisson(),
              control = parglm.control(nthreads = 1L))
```

## Standard errors

The default model-based standard errors assume the Poisson variance equals
the mean. They will be too small here because the cluster random effects
induce overdispersion.

```{r model-based}
coeftest(fit)
```

## Heteroskedasticity-consistent (HC) standard errors

`vcovHC()` computes sandwich standard errors that are robust to
misspecification of the variance function. HC3 (the default) is recommended
for small to moderate samples.

```{r hc}
coeftest(fit, vcov = vcovHC)
```

## Cluster-robust standard errors

`vcovCL()` accounts for within-cluster correlation, which is the appropriate
correction here.

```{r cluster}
coeftest(fit, vcov = vcovCL, cluster = ~cluster)
```

As expected, the cluster-robust standard errors are larger than the
model-based ones, reflecting the extra variability due to the cluster
random effects.

## Covariance matrices directly

The covariance matrices themselves are also available:

```{r vcov}
vcovHC(fit, type = "HC3")
vcovCL(fit, cluster = ~cluster)
```

## Regression tables with gtsummary

The [`gtsummary`](https://cran.r-project.org/package=gtsummary) package
produces publication-ready regression tables from model objects.
`parglm()` models are supported directly because they inherit from `glm`.

```{r gtsummary-setup, eval = requireNamespace("gtsummary", quietly = TRUE) && requireNamespace("broom", quietly = TRUE) && requireNamespace("broom.helpers", quietly = TRUE), message = FALSE}
library(gtsummary)
```

### Logistic regression

We fit a logistic regression model with `parglm()` and pass it to
`tbl_regression()`. Setting `exponentiate = TRUE` displays odds ratios with
their confidence intervals.

```{r gtsummary-logistic, eval = requireNamespace("gtsummary", quietly = TRUE) && requireNamespace("broom", quietly = TRUE) && requireNamespace("broom.helpers", quietly = TRUE)}
set.seed(2)
n2    <- 300
x1_b  <- rnorm(n2)
x2_b  <- rnorm(n2)
y_bin <- rbinom(n2, 1, plogis(0.4 + 0.6 * x1_b - 0.4 * x2_b))
dat_b <- data.frame(y = y_bin, x1 = x1_b, x2 = x2_b)

fit_logistic <- parglm(y ~ x1 + x2, data = dat_b, family = binomial(),
                       control = parglm.control(nthreads = 1L))

suppressWarnings(
  tbl_regression(fit_logistic, exponentiate = TRUE)
)
```

### Robust standard errors in a gtsummary table

`tidy_parglm_robust()` is a drop-in `tidy_fun` for `tbl_regression()` that
replaces the default model-based standard errors with sandwich estimates.
Here we use cluster-robust standard errors for the Poisson model fitted
earlier.

```{r gtsummary-robust, eval = requireNamespace("gtsummary", quietly = TRUE) && requireNamespace("broom", quietly = TRUE) && requireNamespace("broom.helpers", quietly = TRUE)}
tbl_regression(fit, tidy_fun = tidy_parglm_robust)
```

## Note

`model = TRUE` (the default in `parglm`) must be set so that the model frame
is stored, allowing `sandwich` to reconstruct the design matrix internally.
